A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface

### 2021-09-30 sections/appendix/waiting_time_formula/main.tex
Apply changes based on Alex's suggestions
Changes to Methodology and Model overview sections
- Chnages in some notation
- Define expressions only once and then just reference them
- Changes in some labels (including some more sections)
- Other minor rewordings

Now just like in the type 1 individuals case the sojourn time is needed. 
For type 2 individuals whenever individuals are at any row apart from the 
first one they automatically get a wait time of \(0\) since they are essentially 
blocked.

@@

\begin{equation} \label{eq:waiting-time-at-each-state-type-1}
    w^{(1)}(u,v) = 
    \begin{cases} 
        0, \hspace{4.85cm} & \textbf{if } (u,v) \notin S_w \\
        c^{(1)}(u,v) + w^{(1)}(u-1, v), & \textbf{if } u > 0 \textbf{ and } v = T \\
        c^{(1)}(u,v) + w^{(1)}(u, v-1), & \textbf{otherwise}
    \end{cases}
\end{equation}

\begin{equation} \label{eq:waiting-time-at-each-state-type-2}
    w^{(2)}(u,v) = 
    \begin{cases} 
        0, \hspace{4.85cm} & \textbf{if } (u,v) \notin S_w \\
        c^{(2)}(u,v) + w^{(2)}(u-1, v), & \textbf{if } u > 0 \textbf{ and } v = T \\
        c^{(2)}(u,v) + w^{(2)}(u, v-1), & \textbf{otherwise}
    \end{cases}
\end{equation}

@@

\begin{equation} \label{eq:recursive-waiting-time-type-1}
    W^{(1)} = \frac{\sum_{(u,v) \in S_A^{(1)}} w^{(1)}(u,v) 
    \pi_{(u,v)}}{\sum_{(u,v) \in S_A^{(1)}} \pi_{(u,v)}}
\end{equation}

\begin{equation}\label{eq:recursive-waiting-time-type-2}
    W^{(2)} = \frac{\sum_{(u,v) \in S_A^{(2)}} w^{(2)}(u,v) \pi_{(u,v)}}
    {\sum_{(u,v) \in S_A^{(2)}} \pi_{(u,v)}}
\end{equation}

>>>>

Now, like in the type 1 individuals case, the sojourn time is needed. 
For type 2 individuals whenever individuals are at any row apart from the 
first one they automatically get a wait time of \(0\) since they are essentially 
blocked.

@@

\begin{equation} \label{eq:waiting_time_at_each_state_type_1}
    w^{(1)}(u,v) = 
    \begin{cases} 
        0, \hspace{4.85cm} & \textbf{if } (u,v) \notin S_w \\
        c^{(1)}(u,v) + w^{(1)}(u-1, v), & \textbf{if } u > 0 \textbf{ and } v = T \\
        c^{(1)}(u,v) + w^{(1)}(u, v-1), & \textbf{otherwise}
    \end{cases}
\end{equation}

\begin{equation} \label{eq:waiting_time_at_each_state_type_2}
    w^{(2)}(u,v) = 
    \begin{cases} 
        0, \hspace{4.85cm} & \textbf{if } (u,v) \notin S_w \\
        c^{(2)}(u,v) + w^{(2)}(u-1, v), & \textbf{if } u > 0 \textbf{ and } v = T \\
        c^{(2)}(u,v) + w^{(2)}(u, v-1), & \textbf{otherwise}
    \end{cases}
\end{equation}

@@

\begin{equation} \label{eq:recursive_waiting_time_type_1}
    W^{(1)} = \frac{\sum_{(u,v) \in S_A^{(1)}} w^{(1)}(u,v) 
    \pi_{(u,v)}}{\sum_{(u,v) \in S_A^{(1)}} \pi_{(u,v)}}
\end{equation}

\begin{equation}\label{eq:recursive_waiting_time_type_2}
    W^{(2)} = \frac{\sum_{(u,v) \in S_A^{(2)}} w^{(2)}(u,v) \pi_{(u,v)}}
    {\sum_{(u,v) \in S_A^{(2)}} \pi_{(u,v)}}
\end{equation}

### 2021-09-30 sections/healthcare_application/main.tex
Minor changes to other sections
Changes to Methodology and Model overview sections
- Chnages in some notation
- Define expressions only once and then just reference them
- Changes in some labels (including some more sections)
- Other minor rewordings

\begin{itemize}
    \item \( \lambda_2 \): The rate of non-emergency patients that the EMS 
    receives
    \item \( \lambda_{1_i} \): The arrival rate of other patients to ED \(i\in\{1, 2\}\)
    \item \( \mu_i \): The service rate of patients at ED \(i\in\{1, 2\}\)
    \item \( C_i \): The number of available resources (beds) in ED \(i\in\{1, 2\}\)  
    \item \( T_i \): The action that ED \(i\in\{1, 2\} \) chooses to play which 
    corresponds to the threshold at which they do not accept EMS patients.
    \item \( N_i \): The total patient capacity of ED \(i\in\{1, 2\} \)
    \item \( M_i \): The total parking capacity of ED \(i\in\{1, 2\} \)
    \item \( R \): The time target for both EDs
    \item \( \alpha \in [0, 1] \) : \textit{Weighting} of blocking time and 
    lost individuals (equation \ref{eq:obj-distributor})
\end{itemize}

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in \ref{sec:model_overview}, the interaction between the two EDs 
is a Normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect-info-game}). This corresponds to reacting to experienced delays.

@@

The use of a learning algorithm allows to investigate, not only the outcome of 
the game, but also how that outcome is reached.
Consider figure \ref{fig:ard-by-itself}. 
By running asymmetric replicator dynamics on the system the behaviour that 
emerges can be observed.
It can be seen that for this particular set of parameters the strategies of 
the two hospitals converge over time. 
Both hospital 1 (row player) and hospital 2 (column player) seem to be playing 
the same strategy \(s_{10}\) which indicates that thresholds \(T_1 = 10\) and 
\(T_2 = 10\) are played.
What is more important in this example is how the two hospitals reached these
decisions which also highlights the importance of using a learning algorithm.
Hospital 2 is able to reach the decision in a short amount of time while 
hospital 1 takes longer and goes through numerous strategies to get there.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.4, trim=120 0 120 0]{imgs/asymmetric_rd/asymmetric_rd.png}
    \caption{Asymmetric replicator dynamics run}
    \label{fig:ard-by-itself}
\end{figure}

>>>>

\begin{itemize}
    \item \( \lambda_2 \): The rate of non-emergency patients that the EMS 
    receives
    \item \( \lambda_{1_i} \): The arrival rate of other patients to ED \(i\in\{1, 2\}\)
    \item \( \mu_i \): The service rate of patients at ED \(i\in\{1, 2\}\)
    \item \( C_i \): The number of available resources (beds) in ED \(i\in\{1, 2\}\)  
    \item \( T_i \): The action that ED \(i\in\{1, 2\} \) chooses to play which 
    corresponds to the threshold at which they do not accept EMS patients.
    \item \( N_i \): The total patient capacity of ED \(i\in\{1, 2\} \)
    \item \( M_i \): The total parking capacity of ED \(i\in\{1, 2\} \)
    \item \( R \): The time target for both EDs
    \item \( \alpha \in [0, 1] \) : \textit{Weighting} of blocking time and 
    lost individuals (equation \ref{eq:obj_distributor})
\end{itemize}

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in \ref{sec:model_overview}, the interaction between the two EDs 
is a Normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). This corresponds to reacting to experienced delays.

@@

The use of a learning algorithm allows to investigate, not only the outcome of 
the game, but also how that outcome is reached.
Consider figure \ref{fig:ard_by_itself}. 
By running asymmetric replicator dynamics on the system the behaviour that 
emerges can be observed.
It can be seen that for this particular set of parameters the strategies of 
the two hospitals converge over time. 
Both hospital 1 (row player) and hospital 2 (column player) seem to be playing 
the same strategy \(s_{10}\) which indicates that thresholds \(T_1 = 10\) and 
\(T_2 = 10\) are played.
What is more important in this example is how the two hospitals reached these
decisions which also highlights the importance of using a learning algorithm.
Hospital 2 is able to reach the decision in a short amount of time while 
hospital 1 takes longer and goes through numerous strategies to get there.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.4, trim=120 0 120 0]{imgs/asymmetric_rd/asymmetric_rd.png}
    \caption{Asymmetric replicator dynamics run}
    \label{fig:ard_by_itself}
\end{figure}

### 2021-09-30 sections/methodology/main.tex
Changes to Methodology and Model overview sections
- Chnages in some notation
- Define expressions only once and then just reference them
- Changes in some labels (including some more sections)
- Other minor rewordings

Th problem defined in section \ref{sec:model_overview} describes a normal-form
game between the decision makers of two queueing systems and a third player that 
decides how to distribute individuals to the systems.
The strategy space of the two queueing systems is defined as the possible values
that the threshold parameter can take \(T_i \in (1, N_i)\).
Then, the distributor has to decide on the proportion of individuals to send to 
each queueing system \(p_A \text{ and } p_B\), where \(p_A, p_B \in [0, 1] \)
and \(p_A + p_B = 1\).
Figure \ref{fig:imperfect-info-game} shows a diagrammatic 
representation of the game to be played and the decisions to be made.

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[-, node distance = 3cm, scale=0.8]
        \node[anchor=north](HA){\(Q_A\)};
        \node[anchor=north](HA_d1) at (3, 2){.};
        \node[anchor=north](HA_d2) at (3, -2){.};

        \path[->] (HA) edge node {}(HA_d1);
        \path[->] (HA) edge node {}(HA_d2);
        \path (HA_d1) edge [bend left] node {}(HA_d2);
        \path (HA_d1) [dashed] edge node {}(HA_d2);

        \node[anchor=north](HB) at (4.1, 0){\(Q_B\)};
        \node[anchor=north](HB_d1) at (6.9, 2){.};
        \node[anchor=north](HB_d2) at (6.9, -2){.};

        \path[->] (HB) edge node {}(HB_d1);
        \path[->] (HB) edge node {}(HB_d2);
        \path(HB_d1) edge [bend left] node {}(HB_d2);

        \node[anchor=north](D) at (7.8, 0){\(D\)};
        \node[anchor=north](D_d1) at (10.8, 2){.};
        \node[anchor=north](D_d2) at (10.8, -2){.};

        \path[->] (D) edge node {}(D_d1);
        \path[->] (D) edge node {}(D_d2);
        \path(D_d1) edge [bend left] node {}(D_d2);
    \end{tikzpicture}
    \caption*{Figure 2 (revisited): Imperfect information Extensive Form Game 
    between the distributor and the 2 queueing systems}
\end{figure}

As described in section \ref{sec:model_overview} queueing system \(A\) decides
on a strategy, then queueing system \(B\) chooses its own threshold, unaware 
of the first queueing system's choice.
Finally, the distributor makes its choice based on the strategies that the 
queueing systems chose to play.

The problem can be formulated by 3 matrices; the two payoff matrices of the 
normal form game and the routing matrix.
The payoff matrices and their utilities are defined by equations 
(\ref{eq:payoff-entry}) and (\ref{eq:payoff_matrices}) is section 
\ref{sec:model_overview}:

\begin{equation*}
    A = 
    \begin{pmatrix}
        U_{1,1}^A & U_{1,2}^A & \dots & U_{1,N_B}^A \\
        U_{2,1}^A & U_{2,2}^A & \dots & U_{2,N_B}^A \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^A & U_{N_A,2}^A & \dots & U_{N_A,N_B}^A \\
    \end{pmatrix},
    B = 
    \begin{pmatrix}
        U_{1,1}^B & U_{1,2}^B & \dots & U_{1,N_B}^B \\
        U_{2,1}^B & U_{2,2}^B & \dots & U_{2,N_B}^B \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^B & U_{N_A,2}^B & \dots & U_{N_A,N_B}^B \\
    \end{pmatrix}
    \tag{\ref{eq:payoff_matrices} revisited}
\end{equation*}

Similarly, the entries of the routing matrix are defined by equation 
(\ref{eq:obj_distributor}):
% TODO: Make sure this is defined PROPERLY and CONSISTENTLY in model_overview section or here in model_overview section or here
\begin{equation*}
    (p_A, p_B) \quad s.t. \quad 
    \alpha L_A(p_A) + (1 - \alpha) B_A(p_A) = 
    \alpha L_B(p_B) + (1 - \alpha) B_B(p_B)
    \tag{\ref{eq:obj_distributor} revisited}
\end{equation*}

Equation \ref{eq:obj_distributor} is properly defined and explained in section 
\ref{sec:model_overview}.
Thus, using equation \ref{eq:obj_distributor} for all possible sets of 
thresholds we can get the full routing matrix that consists of the proportions
to send to queueing system A (\(p_A\)) and to queueing system B (\(p_B\)).

@@

In order to populate the payoff matrices and the routing matrix the method
of backwards induction is used.
Consider figure \ref{fig:imperfect-info-game} and the flow of the game that was
described (i.e. \(Q_A, Q_B \rightarrow D\)).
Due to the fact that the payoff matrices \(A\) and \(B\) depend on the routing 
matrix \(R\) the entries of the matrices are calculated in a backwards way 
(\(D \rightarrow Q_A, Q_B\)). 
Thus for every pair of strategies \(T_A, T_B\), the proportion of individuals 
that are to be transported to each queueing system is calculated first.

Consider a game where the capacities of the two systems are \(N_A = 4\) and 
\(N_B = 3\).
The strategy space of players \(Q_A\) and \(Q_B\) would then be 
\(T_A = \{1, 2, 3, 4\}\) and \(T_B = \{1, 2, 3\}\) respectively.
Now, starting from an arbitrary starting point of \(T_A=1\) and \(T_B=1\), the
corresponding entry of the routing matrix the routing matrix can be calculated.
Using equation \ref{eq:obj_distributor} for \(T_A=1\) and \(T_B=1\):

\begin{equation*}
    (p_A, p_B) \quad s.t. \quad 
    \alpha L_A(p_A) + (1 - \alpha) B_A(p_A) = 
    \alpha L_B(p_B) + (1 - \alpha) B_B(p_B)
\end{equation*}

The above equation can be solved by using Brent's bisection algorithm.
% TODO: Cite Brent's algorithm
The calculated value of \(p_A\) corresponds to the entry on the first row and 
first column of the routing matrix:

@@

% TODO: Make sure this is defined PROPERLY and CONSISTENTLY in model_overview section or here
\begin{equation*}
    U_{1, 1}^A = \quad 1 -\left( 
        \hat{P} - P(W_A < R) 
    \right)^2,
    \quad
    U_{1, 1}^B = \quad 1 -\left( 
        \hat{P} - P(W_B < R) 
    \right)^2 \tag{\ref{eq:payoff-entry} revisited}
\end{equation*}

These are the utilities of players \(A\) and \(B\) when both players choose a
strategy of \(T = 1\).
% TODO: Change R everywhere to t
Here \(R\) is the predefined waiting time target to be met by a percentage of 
individuals \( \hat{P} \) and \(P(W_i < R)\) is properly defined in section
\ref{sec:proportion_within_target}.

>>>>

Th problem defined in section \ref{sec:model_overview} describes a normal-form
game between the decision makers of two queueing systems and a third player that 
decides how to distribute individuals to the systems.
The strategy space of the two queueing systems is defined as the possible values
that the threshold parameter can take \(T_i \in (1, N_i)\).
Then, the distributor has to decide on the proportion of individuals to send to 
each queueing system \(p_A \text{ and } p_B\), where \(p_A, p_B \in [0, 1] \)
and \(p_A + p_B = 1\).
Figure \ref{fig:imperfect_info_game} form section \ref{sec:model_overview} 
shows a diagrammatic representation of the game to be played and the decisions 
to be made.
As described in section \ref{sec:model_overview} queueing system \(A\) decides
on a strategy, then queueing system \(B\) chooses its own threshold, unaware 
of the first queueing system's choice.
Finally, the distributor makes its choice based on the strategies that the 
queueing systems chose to play.

The problem can be formulated by 3 matrices; the two payoff matrices of the 
normal form game and the routing matrix.
The payoff matrices and their utilities are defined by equations 
(\ref{eq:payoff_entry}) and (\ref{eq:payoff_matrices}) in section 
\ref{sec:model_overview}.

The routing matrix holds the values \((p_A, p_B)\) which are the proportion 
of class 2 individuals to send to to queueing system \(A\) and \(B\).
Each pair \((p_A, p_B)\) can be calculated using equation 
(\ref{eq:obj_distributor}).
Equation \ref{eq:obj_distributor} is defined and explained in section 
\ref{sec:model_overview}.
Thus, using equation \ref{eq:obj_distributor} for all possible sets of 
thresholds we can get the full routing matrix that consists of the proportions
to send to queueing system A (\(p_A\)) and to queueing system B (\(p_B\)).

@@

In order to populate the payoff matrices and the routing matrix the method
of backwards induction is used.
Consider figure \ref{fig:imperfect_info_game} and the flow of the game that was
described (i.e. \(Q_A, Q_B \rightarrow D\)).
Due to the fact that the payoff matrices \(A\) and \(B\) depend on the routing 
matrix \(R\) the entries of the matrices are calculated in a backwards way 
(\(D \rightarrow Q_A, Q_B\)). 
Thus for every pair of strategies \(T_A, T_B\), the proportion of individuals 
that are to be transported to each queueing system is calculated first.

Consider a game where the capacities of the two systems are \(N_A = 4\) and 
\(N_B = 3\).
The strategy space of players \(Q_A\) and \(Q_B\) would then be 
\(T_A = \{1, 2, 3, 4\}\) and \(T_B = \{1, 2, 3\}\) respectively.
Now, starting from an arbitrary starting point of \(T_A=1\) and \(T_B=1\), the
corresponding entry of the routing matrix the routing matrix can be calculated.
Using equation (\ref{eq:obj_distributor}) for \(T_A=1\) and \(T_B=1\):
\begin{align}
    & \alpha L_A(p_A;T_A=1;T_B=1) + (1 - \alpha) B_A(p_A;T_A=1;T_B=1) = 
    \nonumber \\
    & \alpha L_B(p_B;T_A=1;T_B=1) + (1 - \alpha) B_B(p_B;T_A=1;T_B=1) 
    \label{eq:obj_distributor_1_1}
\end{align}

The values of \(p_A\) and \(p_B\) that satisfy equation 
(\ref{eq:obj_distributor_1_1}) can be found by using Brent's bisection 
algorithm.
% TODO: Cite Brent's algorithm
The calculated value of \(p_A\) corresponds to the entry on the first row and 
first column of the routing matrix:

@@

\begin{align}
    U_{1, 1}^A = 1 -\left( 
        \hat{P} - P(W_A(T_A=1, T_B=1) < t) 
    \right)^2 \nonumber
    \\
    U_{1, 1}^B = 1 -\left( 
        \hat{P} - P(W_B(T_A=1, T_B=1) < t) 
    \right)^2 \label{eq:payoff_entry_1_1}
\end{align}

These are the utilities of players \(A\) and \(B\) when both players choose a
strategy of \(T = 1\).
Here \(R\) is the predefined waiting time target to be met by a percentage of 
individuals \( \hat{P} \) and \(P(W_i < t)\) is properly defined in section
\ref{sec:proportion_within_target}.
